```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(lme4)
```
```{r}
BPRS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/BPRS.txt", sep = " ", header = T)
RATS <- read.table("https://raw.githubusercontent.com/KimmoVehkalahti/MABS/master/Examples/data/rats.txt", sep = "\t", header = TRUE)

BPRSL <- read.csv("BPRSL.csv")
RATSL <- read.csv("RATSL.csv")
glimpse(BPRSL)
glimpse(RATSL)
```
#Just loaded the datasets for analysis and glimpsed them.

```{r}
BPRSL$treatment <- factor(BPRSL$treatment)
BPRSL$subject <- factor(BPRSL$subject)
glimpse(BPRSL)

RATSL$ID <- factor(RATSL$ID)
RATSL$Group <- factor(RATSL$Group)
glimpse(RATSL)
```
#I factorized the categorical variables in both datasets for easier analysis.

```{r}
RATSL <- RATSL %>%
  group_by(Time) %>%
  mutate(stdWeight = (Weight - mean(Weight))/sd(Weight) ) %>%
  ungroup()

ggplot(RATSL, aes(x = Time, y = stdWeight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:5, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(name = "Stand. Weight")
```
#I standardized the data for the plot and then plotted the data. Here we can see it divided by groups. There is an outlier in each group and the outlier is a particularly low weight in groups 1 and 3, while in group 2 the outlier has a higher weight. It also looks like most rats gained weight over the course of the study in group 1, while group 3 seems to have a decrease in weight over time. 

```{r}
RATSL10S <- RATSL %>%
  filter(Time > 1) %>%
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()
```
#I created a summary data by diet group and subject(ID) with mean as the summary variable.

```{r}
ggplot(RATSL10S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=20, size=3, fill = "grey") +
  scale_y_continuous(name = "mean(weight), Day 1-64")
```
#Now I plotted the mean summary measures to a boxplot. We can once again see that each group has an outlier and that the weights increased throughout the study. Group 1 contained rats which were much smaller than groups 2 and 3. 

```{r}
BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRS_reg)
```
#Above are the results from fitting a linear regression model to BPRS data with BPRS-points as response variable and treatment and week as explanatory variables. 

```{r}
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE)
summary(BPRS_ref)
```
#And the next step, above the results from fitting a random intercept model to BPRS data with week and treatment as explanatory variables.

```{r}
BPRS_ref1 <- lmer(bprs ~ week + treatment + (week | subject), data = BPRSL, REML = FALSE)
anova(BPRS_ref1, BPRS_ref)
```
#I also created a random slope model and then performed an Anova test on both the models. The P value came out as 0.01342 and it is low, meaning it has a good fit versus the comparison model. 

```{r}
Fitted <- fitted(BPRS_ref1)
BPRSL <- BPRSL %>%
  mutate(Fitted)

ggplot(BPRSL, aes(x = week, y = Fitted, group = subject)) +
  geom_line() +
  scale_x_continuous(name = "Time (weeks)", breaks = seq(0, 8, 2)) +
  scale_y_continuous(name = "Fitted BRPS (points)") +
  theme(legend.position = "top")
```
#Now I created a fitted BPRS model from the interaction model and plotted it. The BPRS score lowered through out the study and the highest starting BPRS score lowered the most throughout the study. 










